---
title: "Effects of Lipid EMF Weighting"
author: "Robert M Flight"
date: "`r Sys.time()`"
output:
  pdf_document:
  toc: yes
---

```{r setup, echo=FALSE, include=FALSE}
here_loc = here::here()
data_loc = file.path(here_loc, "testing_stuff/voting_with_lipids")
library(smirfeTools)
library(ggplot2)
library(dplyr)
emf_classes = import_emf_classifications(file.path(data_loc, "all_emfs_2020-01-27_classified.json"))
theme_set(cowplot::theme_cowplot())
```

## Purpose

Lipid filtering seems to result in a lot more EMFs (likely due to splitting up EMFs that truly belong together), so we decided to try weighting the scores of lipids (and potentially metabolites) higher to see if we can increase their propensity in the final list of assigned EMFs.

For each EMF, we will extract the set of EMFs assigned, and check how many are lipids.



## Lipids Only

We will first investigate the effect on weighting just lipids. 

### Data

```{r data}
# assumes we are running in /mlab/scratch/rmflight/smirfeTools_data

list_files = dir(data_loc, pattern = "rds", full.names = TRUE)

use_files = c(list_files %>%
  grep("lipid_.*.rds", ., value = TRUE) %>%
  grep("metabolite|weights", ., value = TRUE, invert = TRUE),
  list_files %>% grep("nofilter", ., value = TRUE))

filename_which = data.frame(file = use_files,
                            stringsAsFactors = FALSE)
filename_which = filename_which %>% dplyr::mutate(weight = gsub("lipid_", "", basename(file)) %>% gsub(".rds", "", .),
                                                  base_file = basename(file))
filename_which = dplyr::mutate(filename_which, weight2 = dplyr::case_when(
  grepl("nofilter", weight) ~ 1,
  TRUE ~ as.numeric(weight)
))

file_emfs = purrr::map(use_files, function(in_file){
  tmp = readRDS(in_file)
  tmp$emf_info
})
names(file_emfs) = basename(use_files)
```

### Check Number of Sudo EMFs

Ideally, the number of EMFs found didn't change, or at least didn't change by much.

```{r n_emf}
n_emf = purrr::imap_dfr(file_emfs, function(in_emf, in_file){
    dplyr::summarise(in_emf, n_emf = length(unique(sudo_EMF))) %>%
    dplyr::mutate(base_file = in_file)
})
n_emf = dplyr::left_join(filename_which, n_emf, by = "base_file")
n_emf %>% dplyr::select(base_file, weight2, n_emf) %>% dplyr::arrange(weight2)
```

### Check Number of Lipids in Each Sudo EMF

```{r number_lipids}
just_lipids = dplyr::filter(emf_classes, !(grepl("not_lipid", Categories))) %>%
  dplyr::select(isotopologue_EMF) %>% unique(.)
n_lipid = purrr::imap_dfr(file_emfs, function(in_emf, in_file){
  dplyr::group_by(in_emf, sudo_EMF) %>% 
    dplyr::summarise(n_lipid = sum(unique(isotopologue_EMF) %in% just_lipids$isotopologue_EMF),
                     n_iemf = length(unique(isotopologue_EMF))) %>%
    dplyr::mutate(base_file = in_file)
})
n_lipid = dplyr::left_join(n_lipid, filename_which, by = "base_file")
n_lipid = dplyr::mutate(n_lipid, fraction = n_lipid / n_iemf)

split_lipid = split(n_lipid, n_lipid$base_file)
create_binned_fraction = function(in_fraction){
  use_values = in_fraction$fraction
  binned = hist(use_values, breaks = 10, plot = FALSE)
  data.frame(fraction = binned$mids,
             count = binned$counts,
             weight = in_fraction$weight2[1],
             base_file = in_fraction$base_file[1],
             stringsAsFactors = FALSE)
}
n_lipid_fraction = purrr::map_df(split_lipid, create_binned_fraction)

n_lipid_fraction2 = purrr::map_df(split(n_lipid_fraction, n_lipid_fraction$fraction), function(x){
  x$count_frac = x$count / max(x$count)
  x
})
```

Let's look at the bar plots across all different weights.

```{r plot_bar_all}
ggplot(n_lipid, aes(x = fraction)) + geom_histogram(bins = 10) + 
  facet_wrap(~ weight, ncol = 3)
```

Now we will concentrate on just the 0 bar and the 1 bar in the following graphs, plotting number of EMFs with that fraction.

```{r plot_05_95}
ggplot(dplyr::filter(n_lipid_fraction2, fraction %in% c(0.05, 0.95)), aes(x = weight, y = count)) + geom_point() + geom_line() + facet_wrap(~ fraction, ncol = 2, scales = "free_y") + ggtitle(label = "", subtitle = "Number of sudo_EMFs with X fraction of assignments\n as lipids by weighting applied")

ggplot(dplyr::filter(n_lipid_fraction2, fraction %in% c(0.05, 0.95)), aes(x = weight, y = count_frac)) + geom_point() + geom_line() + facet_wrap(~ fraction, ncol = 2, scales = "free_y") + ggtitle(label = "", subtitle = "Number of sudo_EMFs with X fraction of assignments\n as lipids by weighting applied")
```


## Lipids & Metabolights

And then look at weighting both lipids and not_lipid.

### Data

```{r data2}
# assumes we are running in /mlab/scratch/rmflight/smirfeTools_data

list_files = dir(data_loc, pattern = "rds", full.names = TRUE)

use_files = list_files %>%
  grep("lipid_metabolite.*.rds", ., value = TRUE) 

filename_which = data.frame(file = use_files,
                            stringsAsFactors = FALSE)
filename_which = filename_which %>% dplyr::mutate(weight = gsub("lipid_metabolite_", "", basename(file)) %>% gsub(".rds", "", .),
                                                  base_file = basename(file))
filename_which = dplyr::mutate(filename_which, weight2 = dplyr::case_when(
  grepl("nofilter", weight) ~ 1,
  TRUE ~ as.numeric(weight)
))

file_emfs = purrr::map(use_files, function(in_file){
  tmp = readRDS(in_file)
  tmp$emf_info
})
names(file_emfs) = basename(use_files)
```

### Check Number of Sudo EMFs

Ideally, the number of EMFs found didn't change, or at least didn't change by much.

```{r n_emf2}
n_emf = purrr::imap_dfr(file_emfs, function(in_emf, in_file){
    in_emf %>% 
    dplyr::summarise(n_emf = length(unique(sudo_EMF))) %>%
    dplyr::mutate(base_file = in_file)
})
n_emf = dplyr::left_join(filename_which, n_emf, by = "base_file")
n_emf %>% dplyr::select(base_file, weight2, n_emf) %>% dplyr::arrange(weight2)
```

### Check Number of Lipids and Metabolites (Not) in Each Sudo EMF

```{r number_lipids2}
just_not_lipid = dplyr::filter(emf_classes, (grepl("not_lipid", Categories))) %>%
  dplyr::select(isotopologue_EMF) %>% unique(.) %>%
  dplyr::filter(!(isotopologue_EMF %in% just_lipids$isotopologue_EMF))
n_lipid_n_not = purrr::imap_dfr(file_emfs, function(in_emf, in_file){
  dplyr::group_by(in_emf, sudo_EMF) %>% 
    dplyr::summarise(n_lipid = sum(unique(isotopologue_EMF) %in% just_lipids$isotopologue_EMF),
                     n_not = sum(unique(isotopologue_EMF) %in% just_not_lipid$isotopologue_EMF),
                     n_iemf = length(unique(isotopologue_EMF))) %>%
    dplyr::mutate(base_file = in_file)
})
n_lipid_n_not = dplyr::left_join(n_lipid_n_not, filename_which, by = "base_file")
n_lipid_n_not = dplyr::mutate(n_lipid_n_not, lipid_fraction = n_lipid / n_iemf,
                              not_fraction = n_not / n_iemf)

create_binned_fraction2 = function(in_fraction){
  binned_lipid = hist(in_fraction$lipid_fraction, breaks = 10, plot = FALSE)
  binned_not = hist(in_fraction$not_fraction, breaks = 10, plot = FALSE)
  
  data.frame(fraction = binned_lipid$mids,
             count = binned_lipid$counts,
             not = binned_not$counts,
             weight = in_fraction$weight2[1],
             base_file = in_fraction$base_file[1],
             stringsAsFactors = FALSE)
}

n_lipid_not_fraction = purrr::map_df(split(n_lipid_n_not, n_lipid_n_not$base_file), create_binned_fraction2)

n_lipid_not_fraction2 = purrr::map_df(split(n_lipid_not_fraction, n_lipid_not_fraction$fraction), function(x){
  x$count_frac = x$count / max(x$count)
  x$not_frac = x$not / max(x$not)
  x
})

ggplot(dplyr::filter(n_lipid_not_fraction2, fraction %in% c(0.05, 0.95)), aes(x = weight, y = count)) + geom_point() + geom_line() + facet_wrap(~ fraction, ncol = 2, scales = "free_y") + ggtitle(label = "", subtitle = "Number of sudo_EMFs with X fraction of assignments\n as lipids by weighting applied")

ggplot(dplyr::filter(n_lipid_not_fraction2, fraction %in% c(0.05, 0.95)), aes(x = weight, y = not)) + geom_point() + geom_line() + facet_wrap(~ fraction, ncol = 2, scales = "free_y") + ggtitle(label = "", subtitle = "Number of sudo_EMFs with X fraction of assignments\n as metabolites by weighting applied")

lipid_95 = dplyr::filter(n_lipid_fraction2, fraction %in% 0.95) %>% 
  dplyr::select(count, fraction, weight) %>%
  dplyr::mutate(type = "lipid_only")
lipid_not_95 = dplyr::filter(n_lipid_not_fraction2, fraction %in% 0.95) %>%
  dplyr::select(count, fraction, weight) %>%
  dplyr::mutate(type = "lipid_metabolite")

lipid_compare = rbind(lipid_95, lipid_not_95)
ggplot(lipid_compare, aes(x = weight, y = count, color = type)) + geom_point() + 
  geom_line() + labs(subtitle = "Comparison of # of Lipids Weighting Lipids\n or Lipids & Metabolites", caption = "This compares the number of sudo_EMFs that have 100% of their assignments as a lipid\n with different weights applied to either just the lipid or lipid and metabolites weighted scores.") +
  theme(legend.position = c(0.8, 0.1))

```